---
title: "Boston House Price Prediction"
subtitle: 'Notebook 2, continuation of Notebook 1.'
author: 
  name: "Anurima Dey"
date: 28-April-2020
output: 
  html_document:
  df_print: paged
---

## Purpose of the project: 

This notebook contains the non-regression models for the price prediction of Boston Housing.
Do refer to notebook 1, for all the steps, and visualization of the data and corelation among various attibutes. We shall keep the basic steps same i.e. obtaining the data and the cleaning of data remains same. 

Following that we shall proceed directly with the non regression models.

### Obtaining and cleaning the data. 


The data is obtained from kaggle website. <br> The displayed are the __features__ of the dataset.  


1. CRIM     – per capita crime rate by town <br>
2. ZN       – proportion of residential land zoned for lots over 25,000 sq.ft <br>
3. INDUS    – proportion of non-retail business acres per town <br>
4. CHAS     – Charles River dummy variable (1 if tract bounds river; else 0) <br>
5. NOX      – nitric oxides concentration (parts per 10 million) <br>
6. RM       – average number of rooms per dwelling <br>
7. AGE      – proportion of owner-occupied units built prior to 1940 <br>
8. DIS      – weighted distances to five Boston employment centres<br>
9. RAD      – index of accessibility to radial highways <br>
10. TAX      – full-value property-tax rate per $10,000 <br>
11. PTRATIO  – pupil-teacher ratio by town <br>
12. B        – 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town <br>
13. LSTAT    – % lower status of the population <br>
14. MEDV     – Median value of owner-occupied homes in $1000’s

Read the dataset as a dataframe using read.table. After loading and cleaning the data set looks like the following.

```{r}
Bhousing <- read.table("housing.csv", header = FALSE, sep= "")
column_names=c('CRIM', 'ZN', 'INDUS', 'CHAS', 'NOX', 'RM', 
                'AGE', 'DIS', 'RAD', 'TAX', 'PTRATIO', 'B', 'LSTAT', 'MEDV')
names(Bhousing)=column_names
head(Bhousing)

```


```{r}
class(Bhousing)
```
Now, let us check and explore the cleaned data frame containing the housing data. The __Bhousing__ R object is of class ‘data.frame’. <br> The __str()__ function displays the structure of an R dataframe.

#### What does __str()__ show:
1. The data type.
2. Name and data types of each column.
3. Some of the observations.

```{r}
str(Bhousing)
```

To get more detailed statistical information from each column, __summary()__ function can be used. It displays the summary statistics like *minimum value, maximum value, median, mean, and the 1st and 3rd quartile* values for each column in our dataset.<br> Summary also provides information about the missing values, if any is present. We see that there are no missing values in any of our variables.

```{r}
summary(Bhousing)
```

We shall base our prediction mainly on the attribute MEDV (median value of the owner occupied homes) and also on other attributes that are strongly correlated to it. For the given variables and our intuition, we can feel that mnainly, DIS, RAD, LSTAT, NOX, INDUS, CRIM, TAX should have some effect on the variable MEDV. But we shall try to find out all the different types of dependencies in the next section. 


```{r}
library(caret)
Bhousing.scale=cbind(scale(Bhousing[1:13]), Bhousing[14])
set.seed(10)

inTrain <- createDataPartition(y=Bhousing.scale$MEDV, p=0.70, list= F)
Bhousing.training <- Bhousing.scale[inTrain,]
Bhousing.testing <- Bhousing.scale[-inTrain,]
head(Bhousing.training)

```


```{r}
library(caret)
library(randomForest)
control <- trainControl(method='repeatedcv', 
                        number=10, 
                        repeats=3,
                        search = 'random')

#Random generate 15 mtry values with tuneLength = 15

set.seed(1)
rf_random <- train(log(MEDV) ~ .,
                   data = Bhousing.training,
                   method = 'rf',
                   tuneLength  = 15, 
                   trControl = control)
print(rf_random)

```


```{r}
plot(rf_random$finalModel)
```

```{r}
plot(rf_random)
```

```{r}
rf_random$times
```

```{r}
rf_random$finalModel
```


```{r}
set.seed(1)
pred.m2=predict(rf_random, newdata=Bhousing.testing)
actuals_predicts= data.frame(cbind(actuals=Bhousing.testing$MEDV, predicts=exp(pred.m2)))

head(actuals_predicts)

```

```{r}
library(DMwR)
library(grid)
set.seed(10)
regr.eval(actuals_predicts$actuals,actuals_predicts$predicts)
```


```{r}
layout(matrix(c(1,2),2))
densityplot(Bhousing.testing$MEDV)

```

```{r}
densityplot(exp(pred.m2))

```



